---
title: "Quick evaluation Miguel scores, round 2"
output:
  html_notebook: default
  html_document: default
---

### Synopsis

Evaluation fo the second round of scoring Miguel Vázquez did for the DSP.

#### Libraries
```{r libraries,message=FALSE,warning=FALSE}
require(data.table)
require(ggplot2)
require(xlsx)
require(gridExtra)
require(splitstackshape)
```

### Upload data
```{r upload}
dsp_scored_rd1 <- unique(fread("../round1/DarkSpace-v1.txt",sep="\t",header=T,skip=1))
dsp_scored_rd2 <- unique(data.table(read.xlsx2("./DarkSpace_rank2.xls",sheetIndex=1,header=T,colClasses = c("character","numeric","character","integer","character","numeric","numeric"))))
```

### Explore data

I check the distribution of the relevance and the combined score separating IMEx positive and negative publications and comparing the previous iteration with the current one.

#### Relevance score plot
```{r rel_plot}
g1 <- ggplot(dsp_scored_rd1,aes(x=Relevance,fill=IMEX))
g1 <- g1 + geom_histogram(alpha=0.8,position='identity')
g1 <- g1 + xlab("relevance score")
g1 <- g1 + ylab("Number of publications")
g1 <- g1 + ggtitle("Relevance score distribution, round 1")
g1 <- g1 + xlim(0.0,2.0)
g1 <- g1 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

g2 <- ggplot(dsp_scored_rd2,aes(x=Relevance,fill=IMEX))
g2 <- g2 + geom_histogram(alpha=0.8,position='identity')
g2 <- g2 + xlab("relevance score")
g2 <- g2 + ylab("Number of publications")
g2 <- g2 + ggtitle("Relevance score distribution, round 2")
g2 <- g2 + xlim(0.0,2.0)
g2 <- g2 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

grid.arrange(g1, g2, ncol=2)
```

```{r rel_plot_v2}
g1.1 <- ggplot(dsp_scored_rd1,aes(x=Relevance,fill=IMEX))
g1.1 <- g1.1 + geom_density(alpha=0.8,position='identity')
g1.1 <- g1.1 + xlab("relevance score")
g1.1 <- g1.1 + ylab("Number of publications")
g1.1 <- g1.1 + ggtitle("Relevance score distribution, round 1")
g1.1 <- g1.1 + xlim(0.0,2.0)
g1.1 <- g1.1 + ylim(0,5)
g1.1 <- g1.1 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

g2.1 <- ggplot(dsp_scored_rd2,aes(x=Relevance,fill=IMEX))
g2.1 <- g2.1 + geom_density(alpha=0.8,position='identity')
g2.1 <- g2.1 + xlab("relevance score")
g2.1 <- g2.1 + ylab("Number of publications")
g2.1 <- g2.1 + ggtitle("Relevance score distribution, round 2")
g2.1 <- g2.1 + xlim(0.0,2.0)
g2.1 <- g2.1 + ylim(0,5)
g2.1 <- g2.1 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

grid.arrange(g1.1, g2.1, ncol=2)
```

#### Combined score plot
```{r comb_plot}
g1.2 <- ggplot(dsp_scored_rd1,aes(x=`Combined\ score`,fill=IMEX))
g1.2 <- g1.2 + geom_density(alpha=0.8,position='identity')
g1.2 <- g1.2 + xlab("relevance score")
g1.2 <- g1.2 + ylab("Number of publications")
g1.2 <- g1.2 + scale_x_continuous(limits=c(-0.1,10))
g1.2 <- g1.2 + ylim(0,3)
g1.2 <- g1.2 + ggtitle("Combined score density plot, round 1\nNOTE: Axes are truncated at\nx = 0-10 and y = 0-3")
g1.2 <- g1.2 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

g2.2 <- ggplot(dsp_scored_rd2,aes(x=Combined.score,fill=IMEX))
g2.2 <- g2.2 + geom_density(alpha=0.8,position='identity')
g2.2 <- g2.2 + xlab("relevance score")
g2.2 <- g2.2 + ylab("Number of publications")
g2.2 <- g2.2 + scale_x_continuous(limits=c(-0.1,10))
g2.2 <- g2.2 + ylim(0,3)
g2.2 <- g2.2 + ggtitle("Combined score density plot, round 2\nNOTE: Axes are truncated at\nx = 0-10 and y = 0-3")
g2.2 <- g2.2 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

grid.arrange(g1.2, g2.2, ncol=2)
```

No IMEx true values were found over a score of 1. 

### Checking distribution of scores by data source

Judging by manual evaluation (check the [evaluation file](./eval_miguel_rd2.xlsx)), it seems this second verison of the data incorporates mostly data coming from BioGRID. I think we need to examine how the different data sources are contributing to the evaluation score in order to get a better idea on how it works and what would be reasonable cut-offs when used in practice. 

##### Obtain DSP comparison table for reference
I will take the table that does the comparison at the publication level only for this comparison. 
```{r dsp_comp_download}
if(!exists("dsp_pubcomp")){
        setwd("~/Documents/Projects/dsp/darkspaceproject/dsp_comparison/results/")
        dsp_pubcompNames = fread('cat pubcomp_table_final.txt.gz | gunzip | head -n 1')[, colnames(.SD)]
        dsp_pubcomp = fread('cat pubcomp_table_final.txt.gz | gunzip | grep -v "^Day"')
        setnames(dsp_pubcomp, dsp_pubcompNames)
        setwd("~/Documents/Projects/dsp/DarkSpace/manual_evaluation/round2/")
}
```

##### Incorporating & formatting dataset info to ranked lists
```{r ranked_origin}
dsp_scored_rd1$`#ID` <- as.character(dsp_scored_rd1$`#ID`)
dsp_scored_rd1_ori <- unique(merge(dsp_scored_rd1,dsp_pubcomp,by.x="#ID",by.y="pmid",all.x=T,all.y=F))
dsp_scored_rd1_ori_sel <- dsp_scored_rd1_ori[,.(pmid=`#ID`,
                                                Relevance,
                                                IMEX,
                                                Coverage,
                                                Proteins,
                                                `Protein interest`,
                                                `Combined score`,
                                                reactome = gsub("1","reactome",reactome),
                                                tm_epmc = gsub("1","tm_epmc",tm_epmc),
                                                EVEX = gsub("1","EVEX",EVEX),
                                                BioGRID = gsub("1","BioGRID",BioGRID),
                                                GO_IPI = gsub("1","GO_IPI",GO_IPI),
                                                OmniPath_interactions = gsub("1","OmniPath_interactions",OmniPath_interactions),
                                                OmniPath_ptm = gsub("1","OmniPath_ptm",OmniPath_ptm))]

dsp_scored_rd1_ori_long <- reshape(dsp_scored_rd1_ori_sel,direction="long",v.names="origin",varying=c("reactome","tm_epmc","EVEX","BioGRID","GO_IPI","OmniPath_interactions","OmniPath_ptm"))


dsp_scored_rd2$ID <- as.character(dsp_scored_rd2$ID)
dsp_scored_rd2_ori <- unique(merge(dsp_scored_rd2,dsp_pubcomp,by.x="ID",by.y="pmid",all.x=T,all.y=F))
dsp_scored_rd2_ori_sel <- dsp_scored_rd2_ori[,.(pmid=ID,
                                                Relevance,
                                                IMEX,
                                                Coverage,
                                                Proteins,
                                                Protein.interest,
                                                Combined.score,
                                                reactome = gsub("1","reactome",reactome),
                                                tm_epmc = gsub("1","tm_epmc",tm_epmc),
                                                EVEX = gsub("1","EVEX",EVEX),
                                                BioGRID = gsub("1","BioGRID",BioGRID),
                                                GO_IPI = gsub("1","GO_IPI",GO_IPI),
                                                OmniPath_interactions = gsub("1","OmniPath_interactions",OmniPath_interactions),
                                                OmniPath_ptm = gsub("1","OmniPath_ptm",OmniPath_ptm))]

dsp_scored_rd2_ori_long <- reshape(dsp_scored_rd2_ori_sel,direction="long",v.names="origin",varying=c("reactome","tm_epmc","EVEX","BioGRID","GO_IPI","OmniPath_interactions","OmniPath_ptm"))
```

##### Plots origin comparison
```{r plot_origin1}
g1.3 <- ggplot(dsp_scored_rd1_ori_long,aes(x=Relevance,fill=IMEX))
g1.3 <- g1.3 + geom_histogram(alpha=0.8,position='identity')
g1.3 <- g1.3 + xlab("relevance score")
g1.3 <- g1.3 + ylab("Number of publications")
g1.3 <- g1.3 + ggtitle("Relevance score distribution, round 1")
g1.3 <- g1.3 + facet_grid(origin~.,scales="free_y")
g1.3 <- g1.3 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

g2.3 <- ggplot(dsp_scored_rd2_ori_long,aes(x=Relevance,fill=IMEX))
g2.3 <- g2.3 + geom_histogram(alpha=0.8,position='identity')
g2.3 <- g2.3 + xlab("relevance score")
g2.3 <- g2.3 + ylab("Number of publications")
g2.3 <- g2.3 + ggtitle("Relevance score distribution, round 2")
g2.3 <- g2.3 + facet_grid(origin~.,scales="free_y")
g2.3 <- g2.3 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

grid.arrange(g1.3, g2.3, ncol=2)
```

```{r plot_origin2}
g1.4 <- ggplot(dsp_scored_rd1_ori_long,aes(x=Relevance,fill=IMEX))
g1.4 <- g1.4 + geom_density(alpha=0.8,position='identity')
g1.4 <- g1.4 + xlab("relevance score")
g1.4 <- g1.4 + ylab("Number of publications")
g1.4 <- g1.4 + ggtitle("Relevance score distribution, round 1")
g1.4 <- g1.4 + facet_grid(origin~.,scales="free_y")
g1.4 <- g1.4 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

g2.4 <- ggplot(dsp_scored_rd2_ori_long,aes(x=Relevance,fill=IMEX))
g2.4 <- g2.4 + geom_density(alpha=0.8,position='identity')
g2.4 <- g2.4 + xlab("relevance score")
g2.4 <- g2.4 + ylab("Number of publications")
g2.4 <- g2.4 + ggtitle("Relevance score distribution, round 2")
g2.4 <- g2.4 + facet_grid(origin~.,scales="free_y")
g2.4 <- g2.4 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

grid.arrange(g1.4, g2.4, ncol=2)
```

#### Checking protein interest vs relevance score
```{r plot_comb}
g1.5 <- ggplot(dsp_scored_rd1_ori_long,aes(x=Relevance, y=`Protein interest`,color=IMEX))
g1.5 <- g1.5 + geom_point(alpha=0.4,position='identity')
g1.5 <- g1.5 + xlab("relevance score")
g1.5 <- g1.5 + ylab("protein interest")
g1.5 <- g1.5 + ggtitle("Relevance vs protein interest, round 1")
#g1.5 <- g1.5 + facet_grid(origin~.,scales="free_y")
g1.5 <- g1.5 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

g2.5 <- ggplot(dsp_scored_rd2_ori_long,aes(x=Relevance, y=Protein.interest,color=IMEX))
g2.5 <- g2.5 + geom_point(alpha=0.4,position='identity')
g2.5 <- g2.5 + xlab("relevance score")
g2.5 <- g2.5 + ylab("protein interest")
g2.5 <- g2.5 + ggtitle("Relevance vs protein interest, round 2")
#g2.5 <- g2.5 + facet_grid(origin~.,scales="free_y")
g2.5 <- g2.5 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

grid.arrange(g1.5, g2.5, ncol=2)
```
It seems clear more 'interesting' proteins are concentrated in the lower relevance score regions. Let's see the results per data origin. 
```{r plot_comb_ori}
g1.6 <- ggplot(dsp_scored_rd1_ori_long,aes(x=Relevance, y=`Protein interest`,color=IMEX))
g1.6 <- g1.6 + geom_point(alpha=0.4,position='identity')
g1.6 <- g1.6 + xlab("relevance score")
g1.6 <- g1.6 + ylab("protein interest")
g1.6 <- g1.6 + ggtitle("Relevance vs protein interest, round 1")
g1.6 <- g1.6 + facet_grid(origin~.,scales="free_y")
g1.6 <- g1.6 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

g2.6 <- ggplot(dsp_scored_rd2_ori_long,aes(x=Relevance, y=Protein.interest,color=IMEX))
g2.6 <- g2.6 + geom_point(alpha=0.4,position='identity')
g2.6 <- g2.6 + xlab("relevance score")
g2.6 <- g2.6 + ylab("protein interest")
g2.6 <- g2.6 + ggtitle("Relevance vs protein interest, round 2")
g2.6 <- g2.6 + facet_grid(origin~.,scales="free_y")
g2.6 <- g2.6 + theme(plot.title = element_text(),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               panel.background = element_blank())

grid.arrange(g1.6, g2.6, ncol=2)
```

