---
title: "Comparison relevance scores rounds 4 and 5"
output:
  html_document: default
---

### Synopsis

Comparison of thedifferences between rounds 4 and 5 of relevance score, to determine how stocastic effects can influence the results. 

#### Libraries
```{r libraries,message=FALSE,warning=FALSE}
library(data.table)
library(splitstackshape)
library(ggplot2)
```

### Upload data
```{r upload,warning=FALSE,message=FALSE}
dsp_scored_rd4 <- unique(fread("./round4/pmid_ranks.txt",header = T,check.names = T))
dsp_scored_rd5 <- unique(fread("./round5/pmid_ranks.txt",header = T,check.names = T))
```

### Select only relevant fields
```{r select}
rd4 <- unique(dsp_scored_rd4[,.(
        pmid = X.PMID,
        rel_rd4 = Relevance
)])

rd5 <- unique(dsp_scored_rd5[,.(
        pmid = X.PMID,
        rel_rd5 = Relevance
)])
```

### Sorting and comparison
```{r sort_comp}
rd4 <- rd4[order(-rel_rd4),.(
        pmid,
        rel_rd4,
        order_rd4 = .I
)]
rd5 <- rd5[order(-rel_rd5),.(
        pmid,
        rel_rd5,
        order_rd5 = .I
)]

rel_comp <- unique(merge(
        rd4,
        rd5,
        by="pmid",
        all=T
))

rel_comp_long <- melt(
        rel_comp,
        id.vars = "pmid",
        measure.vars = c("rel_rd4","rel_rd5")
)

rel_comp_sort <- rel_comp[order(order_rd4),.(
        pmid,
        rel_rd4,
        rel_rd5,
        diff_rel = rel_rd4 - rel_rd5,
        order_rd4,
        order_rd5,
        diff_order = order_rd4 - order_rd5
)]


```

### Visualize differences
```{r diff_vis_relcomp_1}
g <- ggplot(rel_comp_long,aes(x=value,fill=variable,alpha=0.2)) + 
        geom_histogram(position = "identity",bins = 50) + 
        ggtitle("Relevance score comparison, rounds 4 and 5")
g
```

```{r diff_vis_relcomp_2}
g1 <- ggplot(rel_comp_sort,aes(x=diff_rel)) + 
        geom_histogram(position = "identity",bins = 1000) + 
        ggtitle("Difference in score distribution, rounds 4 and 5")
g1
```

```{r diff_vis_ordercomp}
g2 <- ggplot(rel_comp_sort,aes(x=diff_order)) + 
        geom_histogram(position = "identity",bins = 1000) + 
        ggtitle("Variations in order, rounds 4 and 5")
g2
```

 - Median in order change: `r median(rel_comp_sort$diff_order)`   
 - Standard deviation in order change: `r sd(rel_comp_sort$diff_order)`  
 - Mean order change: `r mean(rel_comp_sort$diff_order)` (as expected from stochastic effect)  
 - Biggest order changes: `r max((rel_comp_sort$diff_order))` (going up in rd5) /  `r min((rel_comp_sort$diff_order))` (going down in rd5)  
 - Number of extreme changes: `r nrow(rel_comp_sort[abs(rel_comp_sort$diff_order) > 0.10*nrow(rel_comp_sort)])` (Extreme changes defined as those in which the order change is bigger than 10% of the total number of records evaluated)  
    
Relevance score seems to be very sensitive to stochastic effects, with a significant amount of records changing ranking position. 

```{r order_change_eff}
g3 <- ggplot(rel_comp_sort,aes(x=order_rd4,y=abs(diff_rel))) +
        geom_point(alpha=0.1) +
        geom_smooth() +
        ggtitle("Score differences between rounds 4 and 5,\nplotted along sorting order for round 4")
g3

```

```{r order_change_eff_2}
g4 <- ggplot(rel_comp_sort[order_rd4<=10000],aes(x=order_rd4,y=abs(diff_rel))) +
        geom_point(alpha=0.1) +
        geom_smooth() +
        ggtitle("Score differences between rounds 4 and 5,\nplotted along sorting order for round 4 (limited to first 10,000 PMIDs)")
g4

```